---
title: Vignette of the oosse package
author: Stijn Hawinkel
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Vignette of the pengls package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates the use of the __oosse__ package for estimating for estimating the out-of-sample R² and its standard error through resampling algorithms.

\setcounter{tocdepth}{5}
\tableofcontents

# Installation instuctions

```{r}
library(oosse)
```


# Illustrative examples

The _R2oosse_ function works with any pair of fitting and prediction functions. Here we illustrate a number of them, but any prediction function implemented in R can be used. The built-in dataset _Brassica_ is used, which contains _rlog_-transformed gene expression measurements for the 1,000 most expressed genes in the _Expr_ slot, as well as 5 outcome phenotypes in the _Pheno_ slot.

```{r loadBrassica}
data(Brassica)
```

## Linear model

The fitting model must accept at least an outcome vector _y_ and a regressor matrix _x_:

```{r linModelFit} 
fitFunLM = function(y, x){lm.fit(y = y, x = cbind(1, x))}
```

The predictive model must accept arguments _mod_ (the fitted model) and _x_, the regressor matrix for a new set of observations.

```{r linModelPredict} 
predFunLM = function(mod, x) {cbind(1,x) %*% mod$coef}
```

Now that these functions have been defined, we apply the prediction model for leaf_8_width using the first 10 genes. Multithreading is used automatically using the _BiocParallel_ package; and an estimate of the computation time is given.

```{r LMpred}
R2lm = R2oosse(y = Brassica$Pheno$Leaf_8_width, x = Brassica$Expr[, 1:10], fitFun = fitFunLM, predFun = predFunLM)
```

Estimates and standard error of the different components are now available.

```{r lmests}
#R2
R2lm$R2
#MSE
R2lm$MSE
#MST
R2lm$MST
```

Also confidence intervals can be constructed:

```{r confintlm}
# R2
buildConfInt(R2lm)
#MSE, 90% confidence interval
buildConfInt(R2lm, what = "MSE", conf = 0.9)
```

By default, cross-validation (CV) is used to estimate the MSE, and nonparametric bootstrapping is used to estimate the correlation between MSE and MST estimators. Other parameters can be supplied though, e.g. for bootstrap .632 estimation of the MSE and jackknife estimation of the correlation:

```{r lmBoot}
R2lm632jn = R2oosse(y = Brassica$Pheno$Leaf_8_width, x = Brassica$Expr[, 1:10], 
                    fitFun = fitFunLM, predFun = predFunLM, methodMSE = "bootstrap",
                    methodCor = "jackknife")
```

## Regularised linear models

For high-dimensional problems, such as the Brassica dataset, 

# Session info

```{r sessionInfo}
sessionInfo()
```

\clearpage

\printbibliography


