---
title: Vignette of the oosse package
author: Stijn Hawinkel
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Vignette of the pengls package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates the use of the __oosse__ package for estimating for estimating the out-of-sample R² and its standard error through resampling algorithms.

\setcounter{tocdepth}{5}
\tableofcontents

# Installation instuctions

# Illustrative examples

The _R2oosse_ function works with any pair of fitting and prediction functions. Here we illustrate a number of them, but any prediction function implemented in R can be used. The built-in dataset _Brassica_ is used, which contains _rlog_-transformed gene expression measurements for the 1,000 most expressed genes in the _BrassicaExpr_ object, as well as 5 outcome phenotypes in the _BrassicaPheno_ object.

## Linear model

The fitting model must accept at least an outcome vector _y_ and a regressor matrix _x_:

```{r linModelFit} 
fitFunLM = function(y, x){lm.fit(y = y, x = cbind(1, x))}
```

The predictive model must accept arguments _mod_ (the fitted model) and _x_, the regressor matrix for a new set of observations.

```{r linModelPredict} 
predFunLM = function(mod, x) {cbind(1,x) %*% mod$coef}
```

# Session info

```{r sessionInfo}
sessionInfo()
```

\clearpage

\printbibliography


